---
title: "individual1"
author: "Brian Bates"
date: "4/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)

rm(list=ls())
```

## Load and Clean Data

The CSV files were saved locally to improve performance

```{r}

acc <- read.csv("accident.csv", stringsAsFactors=FALSE)
str(acc)
```

## Exploratory Analysis

1)  Let's look more at drunk drivers.  Before, we found that the number of accident involving drunk drivers, was 8769.
```{r}
num.drunk.dr <- sum(acc$DRUNK_DR != 0)
total.accidents = dim(acc)[1]

cat(sep="" , "Of the " , total.accidents , " accidents recorded, " , (num.drunk.dr / total.accidents)*100 , "% are fatal accidents involving alcohol.")
```

```{r}
per <- read.csv("person.csv", stringsAsFactors=FALSE)
str(per)
```

Look at a breakdown of accidents by age and sex

```{r}
per$SEX <- factor(per$SEX) #, levels=c("Male", "Female", "NotReported", "Unknown"))
str(per$SEX)

age.dat <- per

age.dat %>% filter(AGE < 100) %>% filter(SEX == 1 | SEX == 2) %>% ggplot(aes(x = AGE, fill=SEX)) + geom_histogram() + labs(title="Accident breakdown by age and sex")
```

Look at a breakdown of accidents involving alcohol by age and sex
```{r}
age.dat %>% filter(AGE < 100) %>% filter(SEX == 1 | SEX == 2) %>% filter(DRINKING == 1) %>% ggplot(aes(x = AGE, fill=SEX)) + geom_histogram() + labs(title="Accidents involving alcohol by age and sex")
```

Let's repeat this, but this time examine drugs
```{r}
age.dat %>% filter(AGE < 100) %>% filter(SEX == 1 | SEX == 2) %>% filter(DRUGS == 1) %>% ggplot(aes(x = AGE, fill=SEX)) + geom_histogram() + labs(title="Accidents involving drugs by age and sex")
```

So far it's pretty obvious that Males outnumber Females in terms of fatal accidents involving drugs and alcohol but in both sexes the dominant age groups are the mid-twenties.


In recent years, penalties for accidents involving work zones has increased.  The data below shows that Construction work zones suffer the most fatalities.
```{r}
# This still doesn't work to get the legend
# acc$WRK_ZONE = factor(acc$WRK_ZONE, levels=c("None", "Construction", "Maintenance", "Construction or Maintenance", "Utility", "Work Zone, Type Unknown", "Not reported"))
acc$WRK_ZONE = factor(acc$WRK_ZONE)
str(acc)

# Look at which type of work zone has the most fatalities
acc %>% filter(YEAR >= 1982 & WRK_ZONE != 0) %>% ggplot(aes(WRK_ZONE)) + geom_bar()
```

## OTHER QUESTIONS
- Who died in the accident - SEAT_POS
- Work zone accidents
- Look at the lower age group and see if the kids were in REST_USE (child safety seat).  Too much filtering due to reporting style changes.
- Airbag deployments and ejections