---
title: "ds202_lab2"
author: "Brian Bates (bkbates), "
date: "2/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Read in the CSV file
choco = read.csv("choco.csv")
```


**Question 1: What is the overall number of chocolate bars rated?**  
The average rating for a choclate bar is 3.19. The total number of chocolate bars rated is 1852.
```{r}
mean(choco$Rating)
dim(choco)[1]
```

**Question 2: How does the number of ratings depend upon the year?**  
There was a trend of increasing number of reviews from 2006-2015, following a sharp decline in 2017.

```{r}
c <- table(choco$Review.Date)
barplot(c, main="Reviews per Year", xlab="Year", ylab="Reviews")
```
<br/>
**Questin 3:  How do ratings compare across different company locations?**
```{r}
t = table(choco$Company.Location)
order(t, decreasing=TRUE)

first = t[58]
first
second = t[19]
second
third = t[8]
third
```

The companies with the most ratings, in decreasing order, are: U.S.A., France, and Canada.  We can run summary statistics across the ratings of each company but a better approach is to visualize it with a boxplot.  

```{r}
# Construct a new data frame holding only the data of the top three companies
df = choco[choco$Company.Location == "U.S.A." | choco$Company.Location == "France" | choco$Company.Location == "Canada", ]

# Refactor the Company.Location.
df$Company.Location = factor(df$Company.Location, levels=c("U.S.A.", "France", "Canada"))

# Boxplot of ratings for each company provides a visual representation of the summary data
boxplot(df$Rating ~ df$Company.Location, main="Ratings of Top Three Locations", xlab="Company Location", ylab="Rating")
```

**Question 4 & 5: Brainstorm and write down an interesting question that can be answer from the dataset. The question should not duplicate the questions above.**
<br/>
As an extra it is interesting to see where the the cocoa percent was for the top three company locations in the context of individual question two.
```{r, fig.width=7, fig.height=7}
m = matrix(data=c(1, 1, 1, 1, 2, 2, 2, 2), nrow=8, ncol = 1, byrow=TRUE)
layout(m)

boxplot(df$Rating ~ df$Company.Location, main="Ratings of Top Three Locations", xlab="Company Location", ylab="Rating")
boxplot(df$Cocoa.Pct ~ df$Company.Location, main="Ratings of Top Three Locations", xlab="Company Location", ylab="Cocoa Percent")
```
<br/>
Another interesting question is to see who are the top companies in the top company location (USA).
```{r}
# All USA data
usa.data = choco[(choco$Company.Location == "U.S.A."), ]

# Top 10 companies in the USA
(highest.levels = levels(reorder(usa.data$Company, usa.data$Rating))[1:10])

```
