---
title: "Lab3"
author: "Brian Bates, Ravi Patel"
date: "2/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)

```

Question 1:
The first plot shows the distribution of degrees and how each degree was portioned into female and male. While the second plot shows the distribution of a individuals financials ccompared to the their degree. 

```{r}
hclean <- HAPPY
str(hclean)
ggplot(hclean, aes(x = DEGREE, fill = SEX)) + geom_bar()
```
```{r}
ggplot(hclean, aes(x = FINRELA, fill = DEGREE)) + geom_bar()
```

Question 2: A mosaic comparing Health, Sex and Degree into determining happiness.
```{r}
library(ggmosaic)
ggplot(hclean) +
  geom_mosaic(aes(x = product(DEGREE), fill = HEALTH)) +
  facet_grid(SEX~HAPPY) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

Question 3: What was learned from the data set was that males and females with higher levels of educaiton are happier. In addition, males and females with better health are happier. 

